---
title: "Assignment 2 - Data Viz and Wrangling"
author: "Eugene Nguyen"
date: "10/6/2021"
output: github_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages
library(tidyverse)
library(skimr)
library(leaflet)
```

## Data Wrangling
You will need to download two datasets from https://github.com/USCbiostats/data-science-data. 
```{r}
# File names
fn1 <- "chs_individual.csv"
fn2 <- "chs_regional.csv"

# Download data

# Individual
if (!file.exists(fn1))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", destfile = fn1)

# Regional
if (!file.exists(fn2))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", destfile = fn2)

# Read into object
ind <- read.csv(fn1)
reg <- read.csv(fn2)
```

The individual and regional CHS datasets in 01_chs. The individual data includes personal and health characteristics of children in 12 communities across Southern California. The regional data include air quality measurements at the community level. Once downloaded, you can merge these datasets using the location variable. Once combined, you will need to do the following:
```{r}
# Merge data using location variable
df <- merge(ind, reg, by = "townname", all.x = TRUE)
```


1. After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.
```{r}
# Count of final dataframe
count(df)

# Count of original individual dataframe
count(ind)
```
The dataframe and the individual dataset both have 1200 rows, there are no duplicates.


In the case of missing values, impute data using the average within the variables “male” and “hispanic.” If you are interested (and feel adventurous) in the theme of Data Imputation, take a look at this paper on “Multiple Imputation” using the Amelia R package here.
```{r}
# View all columns with NA
missing <- colnames(df)[!complete.cases(t(df))]
missing
```

```{r}
# Fill missing

# Mode function
getmode <- function(v) {
   uniqv <- unique(v[!is.na(v)])
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

# Fill missing numeric with average
df$agepft[is.na(df$agepft)]<-mean(df$agepft,na.rm=TRUE)
df$bmi[is.na(df$bmi)]<-mean(df$bmi,na.rm=TRUE)
df$bmi[is.na(df$bmi)]<-mean(df$bmi,na.rm=TRUE)
df$fev[is.na(df$fev)]<-mean(df$fev,na.rm=TRUE)
df$fvc[is.na(df$fvc)]<-mean(df$fvc,na.rm=TRUE)
df$mmef[is.na(df$mmef)]<-mean(df$mmef,na.rm=TRUE)
df$no_24hr[is.na(df$no_24hr)]<-mean(df$no_24hr,na.rm=TRUE)
df$pm2_5_fr[is.na(df$pm2_5_fr)]<-mean(df$pm2_5_fr,na.rm=TRUE)

# Fill missing integer with mode
df$height[is.na(df$height)]<- getmode(df$height)
df$weight[is.na(df$weight)]<- getmode(df$weight)
df$asthma[is.na(df$asthma)]<- getmode(df$asthma)
df$father_asthma[is.na(df$father_asthma)]<- getmode(df$father_asthma)
df$mother_asthma[is.na(df$mother_asthma)]<- getmode(df$mother_asthma)
df$wheeze[is.na(df$wheeze)]<- getmode(df$wheeze)
df$hayfever[is.na(df$hayfever)]<- getmode(df$hayfever)
df$allergy[is.na(df$allergy)]<- getmode(df$allergy)
df$educ_parent[is.na(df$educ_parent)]<- getmode(df$educ_parent)
df$smoke[is.na(df$smoke)]<- getmode(df$smoke)
df$gasstove[is.na(df$gasstove)]<- getmode(df$gasstove)
```

2. Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). 
```{r}
# Obesity Level
df <- 
  df %>%
  mutate(obesity_level = case_when(bmi < 14 ~ "underweight",
                                   bmi >= 14 & bmi < 22 ~ "normal",
                                   bmi >= 22 & bmi < 24 ~ "overweight",
                                   bmi >= 24 ~ "obese"
  ))
```

To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.
```{r}
# Create a crosstab. Use floor to create integer BMI
table(df$obesity_level, floor(df$bmi))
```
It looks like the BMI counts are categorized correctly!

3. Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.
```{r}
# Smoke Gas Exposure
df <- 
  df %>%
  mutate(smoke_gas_exposure = case_when(smoke == 0 & gasstove == 0 ~ "no exposure",
                                        smoke == 1 & gasstove == 0 ~ "smoke only",
                                        smoke == 0 & gasstove == 1 ~ "gas only",
                                        smoke == 1 & gasstove == 1 ~ "smoke & gas",
  ))

# Validate
table(df$smoke_gas_exposure)
```
Most individuals were exposed to gas only followed by no exposure, smoke and gas, and smoke only.

4. Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”
```{r}
# Town
df %>%
  group_by(townname) %>%
  summarise(mean_fev = mean(fev), sd_fev = sd(fev)) %>%
  arrange(desc(mean_fev)) %>%
  knitr::kable()

# Sex
df %>%
  group_by(male) %>%
  summarise(mean_fev = mean(fev), sd_fev = sd(fev)) %>%
  arrange(desc(mean_fev))%>%
  knitr::kable()

# Obesity Level
df %>%
  group_by(obesity_level) %>%
  summarise(mean_fev = mean(fev), sd_fev = sd(fev)) %>%
  arrange(desc(mean_fev))%>%
  knitr::kable()

# Smoke Gas Exposure
df %>%
  group_by(smoke_gas_exposure) %>%
  summarise(mean_fev = mean(fev), sd_fev = sd(fev)) %>%
  arrange(desc(mean_fev))%>%
  knitr::kable()
```
Insights:
- The town with the highest average FEV is Alpine
- Males tend to have higher average FEV than females
- Individuals classified as obese had the highest average FEV
- Interestingly enough, those that were are not exposed to smoke or gas had the highest average FEV

## Looking at the Data (EDA)
The primary questions of interest are: 1. What is the association between BMI and FEV (forced expiratory volume)? 2. What is the association between smoke and gas exposure and FEV? 3. What is the association between PM2.5 exposure and FEV?

Follow the EDA checklist from week 3 and the previous assignment. Be sure to focus on the key variables. Visualization Create the following figures and interpret them. Be sure to include easily understandable axes, titles, and legends.

1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.
```{r}
# Scatterplots
ggplot(data = df, aes(x = bmi, y = fev)) +
  geom_point() +
  geom_smooth(method = "lm", se=FALSE) +
  facet_wrap(~ townname) +
  labs(x = "BMI", y = "FEV", title = "BMI vs. FEV (By Town)")
```

2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r}
# Stacked Histogram
ggplot(data = df, aes(x = fev, fill = obesity_level)) + 
  geom_histogram()
```

3. Barchart of BMI by smoke/gas exposure.
```{r}
# Barchart
ggplot(df, aes(x = obesity_level, y = fev)) + 
    geom_bar(stat="identity")
```

4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.
```{r}
# FEB & BMI
df %>%
  group_by(obesity_level) %>%
  skim(fev)

# FEB & Smoke/Gas Exposure
df %>%
  group_by(smoke_gas_exposure) %>%
  skim(fev)
```

5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
```{r}
# Create average
pm25_avg <- 
  df %>%
  group_by(townname) %>%
  summarise(pm25_avg = mean(pm25_mass),
            lat = mean(lat),
            lon = mean(lon))

# Generating a color palette
pm.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain = pm25_avg$pm25_avg)

pmmap <- leaflet(pm25_avg) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(round(pm25_avg,2), ' C'), color = ~ pm.pal(pm25_avg),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=pm.pal, values=pm25_avg$pm25_avg,
          title='PM2.5 Mass, C', opacity=1)

pmmap
```


6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.
```{r}
# Scatterplot
ggplot(df, aes(x = pm25_mass, y = fev)) +
  geom_point() +
  geom_smooth(method = "lm")
```


