---
title: "Assignment 4 - HPC and SQL"
author: "Eugene Nguyen"
date: "11/16/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(knitr)
library(microbenchmark)
library(parallel)
```

# HPC

## Problem 1: Make sure your code is nice

Rewrite the following R functions to make them faster. It is OK (and recommended) to take a look at Stackoverflow and Google.

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  # YOUR CODE HERE
  rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  # YOUR CODE HERE
  ans <- as.data.table(t(mat))
  ans <- cumsum(ans)
  ans <- t(ans)
  ans
}


# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), check = "equivalent"
)
```

The last argument, check = “equivalent”, is included to make sure that the functions return the same result.

# Problem 2: Make things run faster with parallel computing

The following function allows simulating PI

```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```


In order to get accurate estimates, we can run this function multiple times, with the following code:

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```


Rewrite the previous code using parLapply() to make it run faster. Make sure you set the seed using clusterSetRNGStream():

```{r}
# YOUR CODE HERE
cluster <- makeCluster(4L)
clusterSetRNGStream(cluster,iseed = 1111)
system.time({
  # YOUR CODE HERE
  ans <- unlist(parLapply(cluster, 1:4000, sim_pi, n=1000))
  print(mean(ans))
  # YOUR CODE HERE
})
```


# SQL

Setup a temporary database by running the following chunk

```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

When you write a new chunk, remember to replace the r with sql, connection=con. Some of these questions will reqruire you to use an inner join. Read more about them here https://www.w3schools.com/sql/sql_join_inner.asp


# Question 1

How many many movies is there avaliable in each rating catagory.

```{sql, connection = con}
select rating, count(*) as N
from film
group by rating
order by count(*) desc
```

# Question 2

What is the average replacement cost and rental rate for each rating category.

```{sql, connection = con}
select rating, avg(replacement_cost) as replacement_cost_AVG, avg(rental_rate) as rental_rate_AVG
from film
group by rating
```

# Question 3

Use table film_category together with film to find the how many films there are with each category ID

```{sql, connection = con}
select a.category_id, count(*) as N
from film_category as a
left join film as b on a.film_id = b.film_id
group by a.category_id
```


# Question 4

Incorporate table category into the answer to the previous question to find the name of the most popular category.

```{sql, connection = con}
select a.category_id, c.name, count(*) as N
from film_category as a
left join film as b on a.film_id = b.film_id
left join category as c on a.category_id = c.category_id
group by a.category_id, c.name
order by count(*) desc
limit 1
```

