---
title: "Assignment 03 - Web Scraping and Text Mining"
author: "Eugene Nguyen"
date: "11/5/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Libraries
library(tidyverse)
library(httr)
```

# APIs

* Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine.” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?
```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

# # Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]/span")

# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```
There are 2,336 papers that are found when looking up the keywords "sars-cov-2 trial vaccine".


* Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.
```{r}
### Part 1
library(httr)
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db = "pubmed",
    term = "sars-cov-2 trial vaccine",
    retmax = 1000
  )
)

# Extracting the content of the response of GET
ids <- httr::content(query_ids)



### Part 2
# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "<Id>|</Id>")

# Keep the first 250 IDs
ids <- ids[1:250]

### Part 3
# Create comma separated ids
ids2 = I(paste(ids, collapse = ","))

publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db =  "pubmed",
    id = ids2,
    retmax = 1000,
    rettype = "abstract"
    )
)

# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```


As we did in lab 7. Create a dataset containing the following:

* Pubmed ID number,
* Title of the paper,
* Name of the journal where it was published,
* Publication date, and
* Abstract of the paper (if any).

```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```


```{r}
# Abstract
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>")
abstracts <- str_replace_all(abstracts, "[[:space:]]+", " ")
```



```{r}
# Titles
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
```

```{r}
# Name of Journal
journal <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")
journal <- str_remove_all(journal, "</?[[:alnum:]- =\"]+>")
journal <- str_replace_all(journal, "[[:space:]]+", " ")
```

```{r}
# Date
date <- str_extract(pub_char_list, "<PubDate>[[:print:][:space:]]+</PubDate>")
date <- str_remove_all(date, "</?[[:alnum:]- =\"]+>")
date <- str_replace_all(date, "[[:space:]]+", " ")
```


```{r}
# Create a database
database <- data.frame(
  Journal = journal,
  Date = date,
  PubMedID = ids,
  Title = titles,
  Abstract = abstracts
)
knitr::kable(database[1:20, ], caption = "Abstracts of sars-cov-2 trial vaccine Papers")
```


# Text Mining
A new dataset has been added to the data science data repository https://github.com/USCbiostats/data-science-data/tree/master/03_pubmed. The dataset contains 3241 abstracts from articles across 5 search terms. Your job is to analyse these abstracts to find interesting insights.
```{r}
library(tidytext)
library(tibble)

# Import Data
fn <- "pubmed.csv"

if (!file.exists(fn))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv", destfile = fn)

df <- read.csv(fn)
df <- as.tibble(df)
```


1. Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?
```{r}
# Tokenize the Abstract Column
abstract_20 <- 
  df %>%
  unnest_tokens(token, abstract) %>%
  count(token, sort = TRUE) %>%
  top_n(20, n)

abstract_20 %>%
  knitr::kable()
```

Remove Stop Words
```{r}
# Remove Stop Words
abstract_20_sw <- 
  df %>%
  unnest_tokens(output = word, input = abstract) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  top_n(20)

abstract_20_sw %>%
  knitr::kable()
```

Once stop words were removed, covid, 19, patients, cancer, and prostate were the top words from these abstracts.



2. Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.
```{r}
# Tokenize the Transcription Column into bi-grams/tri-grams
# bi-gram
abstract_bg <- 
  df %>%
  unnest_ngrams(ngram, abstract, n = 2) %>%
  count(ngram, sort = TRUE) %>%
  top_n(10, n)

abstract_bg %>%
  knitr::kable()

# Visualize the top 10 bi-grams
ggplot(abstract_bg, aes(x = ngram, y = n)) +
  geom_bar(stat = "identity") +
  coord_flip()
```


3. Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?
```{r}
df %>%
  unnest_tokens(abstract, abstract) %>%
  count(abstract, term) %>%
  bind_tf_idf(abstract, term, n) %>%
  arrange(desc(tf_idf))
```
COVID is still the top word when compared to Question 1, but now the TF-IDF results show prostate, eclampsia, preeclampsia, and meningitis appear to have the highest values.


